Supplementary Material for Food-web complexity alters the fitness landscape of an insect herbivore

### 2018-06-06 supp_mat_complexity_selection.Rmd
working on visualizations of fitness landscapes.

inverse_logit <- function(x) exp(x)/(1+exp(x))
```

@@

@@

## Control treatment
Wbar_model_control <-  glmer(gall_survival ~ 1 + (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = control_df, glmerControl(optimizer="bobyqa"))
Wbar_control <- inverse_logit(fixef(Wbar_model_control))

glmer_control <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = control_df, glmerControl(optimizer="bobyqa"))
summary(glmer_control)
```

@@

## Ectoparasitoid exclusion treatment
Wbar_model_treatment <-  glmer(gall_survival ~ 1 + (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = treatment_df, glmerControl(optimizer="bobyqa"))
Wbar_treatment <- inverse_logit(fixef(Wbar_model_treatment))

glmer_treatment <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference + 
                           (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = treatment_df, glmerControl(optimizer="bobyqa"))
summary(glmer_treatment)
```

>>>>

inverse_logit <- function(x) exp(x)/(1+exp(x))

Beta_avg_grad <- function(logistic_model, mean_fitness) {
  brackets <- predict(logistic_model, type = "response") * (1 - predict(logistic_model, type="response"))
  gradient <- mean(brackets) * fixef(logistic_model) / mean_fitness 
  return(gradient)
}
```

@@

```{r}
## Calculate directional selection gradients
# following method of Janzen and Stern

Beta_avg_grad <- function(.) {
  mean_fitness <- inverse_logit(fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  brackets <- predict(., type = "response") * (1 - predict(., type="response"))
  mean(brackets) * fixef(.)[-1] / mean_fitness 
}

```

@@

## Control treatment

# GLMM
glmer_control <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference +
                           (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = control_df, glmerControl(optimizer="bobyqa"))
summary(glmer_control)

# Mean fitness
Wbar_control <- inverse_logit(fixef(update(glmer_control, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

# Use parametric bootstrapping to get confidence intervals for selection gradients
boot_Beta_control <- bootMer(glmer_control, FUN = Beta_avg_grad, nsim=100, parallel="multicore", ncpus=32)

data.frame(Beta_avg_grad = round(Beta_avg_grad(glmer_control),2),
           lowerCI_2.5 = round(apply(boot_Beta_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),2),
           upperCI_97.5 = round(apply(boot_Beta_control$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),2)) %>%
  knitr::kable(.)
```

@@

## Ectoparasitoid exclusion treatment

# GLMM
glmer_treatment <- glmer(gall_survival ~ sc.gall_size + sc.clutch_size + sc.female_preference + 
                           (1|Genotype/Plant_Position/Gall_Number),
                         family=binomial(link=logit),
                         data = treatment_df, glmerControl(optimizer="bobyqa"))
summary(glmer_treatment)

# Mean fitness
Wbar_treatment <- inverse_logit(fixef(update(glmer_treatment, .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))

# Use parametric bootstrapping to get confidence intervals for selection gradients
boot_Beta_treatment <- bootMer(glmer_treatment, FUN = Beta_avg_grad, nsim=10)

data.frame(Beta_avg_grad = round(Beta_avg_grad(glmer_treatment),2),
           lowerCI_2.5 = round(apply(boot_Beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.025)),2),
           upperCI_97.5 = round(apply(boot_Beta_treatment$t, 2, FUN = function(bootstraps) quantile(bootstraps, 0.975)),2)) %>%
  knitr::kable(.)
```

```{r}
## Custom function to calculate relative fitness intervals for my models
relative_fitness_intervals <- function(logistic_model, newdata, bootstraps){
  model_matrix <- model.matrix(~ sc.gall_size + sc.clutch_size + sc.female_preference, newdata)
  model_null <- model.matrix(~1, newdata)

  get_relative_fitness <- function(.) {
  #mean_fitness <- mean(inverse_logit(model_null %*% fixef(.))) #fixef(update(., .~. -sc.gall_size -sc.clutch_size -sc.female_preference)))
  absolute_fitness <- inverse_logit(model_matrix %*% fixef(.))
  absolute_fitness #/ mean_fitness # mean(absolute_fitness) # 
  }

  get_bootstraps <-bootMer(logistic_model, FUN = get_relative_fitness, nsim=bootstraps, parallel="multicore", ncpus=32)
  get_intervals <- apply(get_bootstraps$t, 2, function(x) x[order(x)][c(round(bootstraps*0.025,0),round(bootstraps*0.975,0))])

  mean_fitness <- mean(get_relative_fitness(logistic_model))

  return(data.frame(relative_fitness = get_relative_fitness(logistic_model)/mean_fitness,
                    lower_2.5 = get_intervals[1, ]/mean_fitness,
                    upper_97.5 = get_intervals[2, ]/mean_fitness,
                    newdata))
}
```

Illustrating the relationship between phenotype and mean fitness depends critically on the range of parameters shown. Perhaps this is what it is a good idea to show it based on the actual phenotypic distributions, or for 1 +/- SD

```{r}
RF_treatment_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = seq(min(treatment_df$sc.gall_size),
                        max(treatment_df$sc.gall_size),
                        length.out = 100),
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=200)

mean(RF_treatment_gall_size$relative_fitness)

RF_control_gall_size <- relative_fitness_intervals(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = seq(min(control_df$sc.gall_size),
                        max(control_df$sc.gall_size),
                        length.out = 100),
     sc.clutch_size = 0,
     sc.female_preference = 0),
  bootstraps=200)

mean(RF_control_gall_size$relative_fitness)

bind_rows(mutate(RF_treatment_gall_size, Treatment = "Simple"),
          mutate(RF_control_gall_size, Treatment = "Complex")) %>%
  ggplot(., aes(x=sc.gall_size, y=relative_fitness, color=Treatment, fill=Treatment)) +
  geom_hline(yintercept=1, linetype="dotted") +
  geom_line() +
  geom_ribbon(aes(ymin=lower_2.5, ymax=upper_97.5), alpha=0.5)
```

```{r}
RF_treatment_clutch_size <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = seq(min(treatment_df$sc.clutch_size),
                          max(treatment_df$sc.clutch_size),
                          length.out = 100),
     sc.female_preference = 0),
  bootstraps=200)

mean(RF_treatment_clutch_size$relative_fitness)

RF_control_clutch_size <- relative_fitness_intervals(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = seq(min(control_df$sc.clutch_size),
                        max(control_df$sc.clutch_size),
                        length.out = 100),
     sc.female_preference = 0),
  bootstraps=200)

mean(RF_control_clutch_size$relative_fitness)

bind_rows(mutate(RF_treatment_clutch_size, Treatment = "Simple"),
          mutate(RF_control_clutch_size, Treatment = "Complex")) %>%
  ggplot(., aes(x=sc.clutch_size, y=relative_fitness, color=Treatment, fill=Treatment)) +
  geom_hline(yintercept=1, linetype="dotted") +
  geom_line() +
  geom_ribbon(aes(ymin=lower_2.5, ymax=upper_97.5), alpha=0.5)
```

```{r}
RF_treatment_female_preference <- relative_fitness_intervals(
  logistic_model = glmer_treatment, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = 0,
     sc.female_preference = seq(min(treatment_df$sc.female_preference),
                                max(treatment_df$sc.female_preference),
                                length.out = 100)),
  bootstraps=200)

mean(RF_treatment_female_preference$relative_fitness)

RF_control_female_preference <- relative_fitness_intervals(
  logistic_model = glmer_control, 
  newdata = expand.grid(
     sc.gall_size = 0,
     sc.clutch_size = 0,
     sc.female_preference = seq(min(control_df$sc.female_preference),
                                max(control_df$sc.female_preference),
                                length.out = 100)),
  bootstraps=200)

mean(RF_control_female_preference$relative_fitness)

bind_rows(mutate(RF_treatment_female_preference, Treatment = "Simple"),
          mutate(RF_control_female_preference, Treatment = "Complex")) %>%
  ggplot(., aes(x=sc.female_preference, y=relative_fitness, color=Treatment, fill=Treatment)) +
  geom_hline(yintercept=1, linetype="dotted") +
  geom_line() +
  geom_ribbon(aes(ymin=lower_2.5, ymax=upper_97.5), alpha=0.5) +
```
